Red Wine Properties by Micah Shanks
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(GGally)
library(gridExtra)
library(memisc)
library(polycor)
library(RColorBrewer)
```

If running locally, download dataset and insert file location below. 
```{r echo=FALSE, Load_the_Data}
# Load the Data
wine = read.csv("data")
```

# Univariate Plots Section
Below are the names of all initial variables in the dataset, general 
information on the dataset and all variables. In addition a summary is 
produced that shows basic statistics on each variables. Although quality 
was accessed on a scale of 1-10, it is important to note that all wines 
in this dataset were rated between 3-8, with a mean of 5.636. 
Since quality is an ordinal value, mode might be a more accurate 
representation of quality which is 5. 

# Number of columns and rows
The dataset containts 1599 wines, marked by number in the first column of the dataset, followed by 11 measured properties, and a quality rating. 
```{r echo=FALSE}
dim(wine)
```
# Names of the 13 variables
```{r echo=FALSE}
names(wine)
```
# Variable types
```{r echo=FALSE}
str(wine)
```
# Summary of each variable
```{r echo=FALSE}
summary(wine)
```
Calculating the mode for quality
```{r echo=FALSE}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

Mode(wine$quality)
```

# I plotted all variables in the wine dataset below to get an idea of their
# distributions. I changed banwidths on most of the plots, and the scale on 
# some so that more detail could be seen. 

```{r echo=FALSE,warning= FALSE, message=FALSE}
qplot(fixed.acidity, data = wine, binwidth = .10)
shapiro.test(wine$fixed.acidity)
```

There appear to be a few wines that have higher fixed acidity than others, 
but for the most part this data appears normally distributed. 
However, the mean is 8.32, a bit higher than the median of 7.90, 
suggesting it is slightly skewed to the right. 
A Shapiro Wilks test fails to reject the nullhypothesis that the data is normal.  

```{r echo=FALSE,warning= FALSE, message=FALSE}
qplot(volatile.acidity, data = wine, binwidth = .02)
```

The volatile acidity looks to be quite consistent. I wouldn't say the data is
normally distributed as I see two, or maybe arguably three peaks in this plot, 
however the mean and median are very close at .5200 and .5278 respectively. 
There is also an outlier near 1.6, that is quite far from all other measurements.   

```{r echo=FALSE,warning= FALSE, message=FALSE}
qplot(citric.acid, data = wine, binwidth = .02)

```

This data appears to be skewed to the right, however it has a couple of 
interesting aspects to it. It looks like there are many entries that are 
0.0, .25, or .50 in the dataset (shown by the three bars that elevate above
all the rest of the data). This could be because of the way that citric acid
is recorded when measured. Interestingly, although the data is continuous, that 
aspect gives it a ordinal quality. Again I there is an outlier far from others, 
this time at 1.00. 

Because of the extreme outliers in these two plots I took at look at their 
boxplots and found the wines with the associated values. 
Neither are from the same wine. 

```{r echo=FALSE,warning= FALSE, message=FALSE}
volatilebox <- boxplot(wine$volatile.acidity)
volatilebox$out
which(wine$volatile.acidity == 1.580)
```

```{r echo=FALSE,warning= FALSE, message=FALSE}
residualbox <- boxplot(wine$citric.acid)
residualbox$out
which(wine$citric.acid == 1)
```

### Because of outliers in the data, I decided to plot the residual sugar and 
### chlorides on a logarithmic scale in order to take a look at the data in a 
### normal distribution. 

```{r echo=FALSE,warning= FALSE, message=FALSE}
qplot(residual.sugar, data = wine)

qplot(residual.sugar, data = wine, binwidth = .01) +
    scale_x_log10(breaks = c(1:10))
```

The log scale shows that residual sugar in red wine leans toward a normally 
distribution and residual sugar is usually consistent, other than a large 
amount of outliers from the normal distribution. The mean is pulled to 
2.539 by the outliers, while the median is 2.200. 

```{r echo=FALSE, warning= FALSE, message=FALSE}
boxplot(wine$residual.sugar)
```

The boxplot visualizes the mass amount of outliers from the otherwise 
normal distribution.


```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(chlorides, data = wine)

qplot(chlorides, data = wine, binwidth = .02) +
    scale_x_log10(breaks = c(.01, .05, .10, .15, .20))
```

The log scale on sulfur dioxide shows a much clearer picture of the data, 
showing that most of the data lies between .05 and .15, with a fair amount 
of data above .15 that are not included in the normal distribuation of data. 
The median is at .07900 and the mean is at .08747, pulled to the right by
the outliers. 

```{r echo=FALSE, warning= FALSE, message=FALSE}
boxplot(wine$chlorides)
```

Like residual sugar, the boxplot demonstrates the large amount of outliers. 

```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(free.sulfur.dioxide, data = wine)
which.max(wine$free.sulfur.dioxide)
```

Free sulfur dioxide is skewed to the right. The mean is far from the mode of 
38.00 at 46.47. There is a max of 289, again a different wine from others I 
looked up with outliers thus far. 

```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(total.sulfur.dioxide, data = wine) +
  xlim(0, 200)
which.max(wine$total.sulfur.dioxide)
```

Total sulfur dioxide is also skewed to the right. The mean is 46.47, median 
is 38.00 and the max is 289.00 showing a similar distribuation to free sulfur 
dioxides. The wines with the max values however are different.   

```{r echo=FALSE, warning= FALSE, message=FALSE}
cor.test(wine$free.sulfur.dioxide, wine$total.sulfur.dioxide, 
         method = "pearson")
```

Because the plots look similar, I ran a correlation test and found a value of 
.668 which shows a moderate correlation.

```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(density, data = wine)
```

Density appears to be normally distributed. The median and mean are similar 
at .9968 and .9967 respectively. 

```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(pH, data = wine)
```

pH appears normally distributed, with most wines falling inbetween 3.0 and 3.5, 
with an outlier at 4.010 shown by the max. The median and mean are close though
at 3.310 and 3.311.  

```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(sulphates, data = wine, binwidth = .02) +
  xlim(0, 1.5)
```

Sulphates shows slight skew to the right, but for the most part looks normally
distributed with almost all data falling inbetween .5 and 1.0.The mean is 
slightly higher at .6581, compared to the median at .6200.  

```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(alcohol, data = wine, binwidth = .2)

```

Most wines in this dataset fall between 9 and 10 percent alcohol by volume, 
steadily tapering off to around 13 percent. Wines that have under 9 percent
or over 13 percent alcohol are rarer in this dataset. The mean is at 10.42
with the meadian just a bit lower at 10.20.


```{r echo=FALSE, warning= FALSE, message=FALSE}
qplot(quality, data = wine)
```

The plot below shows that most wines in this dataset are rated as 5 or 6 in
quality. That leads me to wonder what correlations can be found that are 
associated with lower (3/4) and higher (7/8) ratings. Are there properties
of wine that are predictors of lower and higher rated wine? 

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 wines in the dataset with 13 variables. One variable, X, 
assigns the wine a number. Quality is a rating, with a minimum of 3 and max
of 8 in this dataset (rated on a scale of 0-10, very bad - very excellent). 
The other 11 variables are chemical properties of the wine.  

### What is/are the main feature(s) of interest in your dataset?
I think the main variable of interest in the dataset is quality. The question
of interest is finding if any of the chemical properties can be used as strong
predictors of the quality of a wine. Because the taste of wine is complex, and
a balance of properties is important, I assume I will be looking for a 
combination of properties that affect quality. My initial interest is
in fixed acidity, residual sugar, and alcohol. 

### What other features in the dataset do you think will help support your 
### investigation into your feature(s) of interest? 
Since most wines fall into quality ratings of 5 and 6 and thus we could call
these average, it is possible that cutting the data set into low ratings (3/4)
and high ratings (7/8) could be a way to look for properties that stand out. 

### Did you create any new variables from existing variables in the dataset?
Bouncing off the idea above I created a variable that groups the ratings in 
low, average, and high.

```{r echo=FALSE}
wineQuality <- function(x) {
    if(is.na(x)){
      NA
    }else if(x == "3"){
     "low"
    }else if(x == "4"){
     "low"
    }else if(x == "5"){
      "average"
    }else if(x == "6"){
      "average"
    }else{
      "high"
    }
}

# Create New Wine Rating Variable
wine$quality.group <-   apply(wine['quality'],1,wineQuality)
```

### Of the features you investigated, were there any unusual distributions? 
### Did you perform any operations on the data to tidy, adjust, or change 
### the form of the data? If so, why did you do this?
I thought the residual sugar, chlorides, and free sulfur dioxide plots were
all being effected by long tails. I performed a log function to adjust the 
plot closer to a normal distribution. After adjusting binwidths as well on all
the plots, it looked like most chemical properties have a normal distribution 
or are skewed to the right. Citric acid is the only property that looks 
sporatic. The distribution is skewed to the right overall, but many values are
at 0.00, 0.25, or .50. This may be a product of the way citric acid is recorded
in many wines. 

I ran a few correlation test on variables that I thought might be correlated 
with the quality rating of wine. I found alcohol to have the highest 
correlation at .47. Surprisingly, residual sugar has effectively no correlation
to the quality ratings of wine. Volatile acidity is negatively correlated with 
quality at a value of -0.39. Some quick research showed me that excessive
volatile acids are associated with unpleasing aromos and poor taste. I also 
found that acids are an important part of the overall character of wine so I 
want to further explore citric acid which shows a correlation with quality of
.22. I also found it interesting that sulphates produced a correlation of .25
since to my understanding this is a perservative used to stop the fermentation
process. 



###Below I produced the Pearson product-moment correlation coefficient test
### among variables of interest and produced a table as well across all variables
### of interest. 

```{r echo=FALSE}
cor.test(as.numeric(wine$quality), wine$fixed.acidity, method = "pearson")
cor.test(as.numeric(wine$quality), wine$volatile.acidity, method = "pearson")
cor.test(as.numeric(wine$quality), wine$citric.acid, method = "pearson")
cor.test(as.numeric(wine$quality), wine$residual.sugar, method = "pearson")
cor.test(as.numeric(wine$quality), wine$alcohol, method = "pearson")
cor.test(as.numeric(wine$quality), wine$sulphates, method = "pearson")
```

```{r echo=FALSE}
corMnames <- c("quality", "fixed.acidity", "volatile.acidity", "citric.acid", 
               "residual.sugar", "alcohol", "sulphates")

corM <- wine[corMnames]

hetcor(corM)
```

```
Correlation matrix, not including in final report

theme_set(them_minimal(20))

set.seed(500)
wine_subset <- wine[, c(2:13)]
names(wine_subset)
ggpairs(wine_subset[sample.int(nrow(wine_subset), 1000),])
```

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
ggplot(aes(x = factor(quality), y = alcohol), data = wine) +
  geom_boxplot()
```

I first ran a scatterplot on quality vs. alcohol, but I found a boxplot is a
more useful visualization of the relationship. It is clear that higher rated
wines have more alcohol. 


```{r echo=FALSE}
wine$quality.group <- factor(wine$quality.group, 
  levels = c("low", "average", "high"))

ggplot(aes(x = quality.group, y = alcohol), data = wine) +
  geom_boxplot()
```

The grouped variable I created gives an even clearer picture. The low and 
average buckets have similar distributions, however the highly rated wines 
have higher alcohol, with the IQR of wines rated high almost completely outside
of the range of wines rated as low and average. It should be noted though that 
there are a few outliers in the average group with high percentages of alcohol. 
At a 95% confidence level, the Pearson correlation coefficient between quality
and alcohol is 0.4761663


```{r echo=FALSE}
ggplot(aes(x = factor(quality), y = residual.sugar), data = wine) +
  geom_boxplot()
```

The boxplot below shows a visualization of residual sugars lack of effect
on quality ratings. The outliers in the quality ratings display the variability
of sugar levels within quality ratings. Along with a Pearson correlcation value
of 0.0137, I am concluding that residual sugar levels, within an expected
(or normal) range do not have a meaningful effect on quality ratings in this
dataset. Pearson correlation coefficient between quality and alcohol is
0.01373164

```{r echo=FALSE}
ggplot(aes(x = factor(quality), y = residual.sugar), data = wine) +
  geom_boxplot() + scale_y_continuous(limits = c (0,4))
```

Removing the outliers gives a clearer picture of the distributions and shows
that the medians are similar across all quality rating. 

 

```{r echo=FALSE}
p1 <- ggplot(aes(x = quality.group, y = alcohol), data = wine) +
  geom_boxplot()

p2 <- ggplot(aes(x = quality.group, y = citric.acid), data = wine)+
  geom_boxplot()

p3 <- ggplot(aes(x = quality.group, y = volatile.acidity), data = wine)+
  geom_boxplot()

p4 <- ggplot(aes(x = quality.group, y = sulphates), data = wine)+
  geom_boxplot()

grid.arrange(p1, p2, p3, p4, ncol=2)
```

This is a group of boxplots of the four variables that I am most interested
in thus far, based on correlations. They all visualize notable trends.

 

```{r echo=FALSE}
p1 <- ggplot(aes(x = factor(quality), y = alcohol), data = wine) +
  geom_boxplot()

p2 <- ggplot(aes(x = factor(quality), y = citric.acid), data = wine)+
  geom_boxplot()

p3 <- ggplot(aes(x = factor(quality), y = volatile.acidity), data = wine)+
  geom_boxplot()

p4 <- ggplot(aes(x = factor(quality), y = sulphates), data = wine)+
  geom_boxplot()

grid.arrange(p1, p2, p3, p4, ncol=2)
```


Here is the same group of plots, except using the original ratings. I think for
the most part, my grouped ratings tell a story that is easier to see. It is
interesting that the clearest trend that emerges from these plots is the 
relationship betweeen volatile acidity and quality.

```{r echo=FALSE}
ggplot(aes(x = factor(quality), y = sulphates), data = wine)+
  geom_boxplot() + scale_y_continuous(limits = c(0, 1.0))
```

Because of the large amount of outliers, truncating the y-axis gives a clearer
picture of the trend between suplphates and quality. 


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the
### investigation. How did the feature(s) of interest vary with other features
### in the dataset?

### &

### Did you observe any interesting relationships between the other features
### (not the main feature(s) of interest)?
A bivariate analysis developed my features of interest. Alcohol made the list
with it's strong effect on quality. where I originally thought fixed.acid
would be an interesting factor (based soley on my lack of expertise), I now 
see that citric acid is a better predictor of quality. I found that sugar does
not have the effect I thought it might, however volatile acidity and sulphate 
levels proved to be chemical properties that seem to have an effect on quality
ratings in this dataset. 

### What was the strongest relationship you found?
Alcohol is the strongest relationship I found. 


```{r echo=FALSE}
wine$alcohol.group <- cut(wine$alcohol, breaks = 3, labels = c("low", "medium", "high"))
```

summary(wine$alcohol)
# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x = quality, y = volatile.acidity), 
       data = wine) +
       geom_point(aes(color = alcohol.group, shape = alcohol.group, size = 1), 
       stat = 'summary', fun.y = median) 
```

I created a variable that defines the alcohol in buckets of "low", "medium",
or "high". This plot shows the median value volatile acidity within each quality
by alcohol group. Here, with quality levels of 5 and 6, the alcohol groups have
similar volatile acidity, where there is more variance at other levels. Among
the lowest quality rating charted (3), the volatile acidity is much lower than
that of the highest rating charted (8). 

```{r echo=FALSE}
ggplot(aes(x = quality, y = citric.acid), 
       data = wine) +
       geom_point(aes(color = alcohol.group, shape = alcohol.group, size = 1), 
       stat = 'summary', fun.y = median) 
```

This chart shows the median suplphate measurements across quality ratings by
alcohol groups. 
The chart shows a trend that is easier to see than the previous where
regardless of alcohol group, the quality rating tends to increase as the
amount of citric acid increases. Especially among medium abv wines, the
relationship looks linear. There is also a trend among groups, where the
highest abv has lowest citric acid measurements among quality groups 5-8. 

```{r echo=FALSE}
ggplot(aes(x = quality, y = sulphates), 
       data = wine) +
       geom_point(aes(color = alcohol.group, shape = alcohol.group, size = 1), 
       stat = 'summary', fun.y = median)

cor.test(wine$alcohol, wine$sulphates, method = "pearson")
```

This chart shows the median sulphate levels across quality ratings by alcohol
groups.
The grouping among quality ratings 4-7 pops out first. There appears to be a
linear relationship and the alcohol groups are have similar sulphate
measurements across the ratings. The lowest and highest ratings (3-8),
have more variance, especially at rating 8. 

```{r echo=FALSE}
wine$quality <- ordered(wine$quality, levels = c(3, 4, 5, 6, 7, 8))

myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality)
colScale <- scale_colour_manual(name = "quality",values = myColors)

ggplot(aes(x = alcohol, y = sulphates),
       data = wine) +
       geom_point(aes(color = quality)) +
       colScale
```

This chart plots each wine based on their alcohol and sulphate levels, colored
by quality rating. Since there isn't a lot of data on the highest and lowest
ratings, it is difficult to see any trends among those ratings. Of the lower
quality wines, I can see points despersed across alcohol levels, at lower
sulphate measurements. There is a grouping of those wines with quality level
5 trending toward lower sulphate and alcohol levels. It seems that wines rated
at quality 6 and 7 tend to have higher alcohol levels, yet are dispersed across
sulphate levels, mostly within the range of .5 and 1.0.

```{r echo=FALSE}
myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality)
colScale <- scale_colour_manual(name = "quality",values = myColors)

ggplot(aes(x = alcohol, y = volatile.acidity),
       data = wine) +
       geom_point(aes(color = quality)) +
       colScale
```

Here I changed the y axis to volatile acidity. I would have expected to see
less dispersion across volatile acidity here across quality ratings. Again
those wines rated 5 are tending toward lower abv. I do see lower quality wines
plotted at higher volatile acidity and the highest quality wines plotted at
lower volatile acidity, however average rated wines show a lot of variance
among volatile acidity measurements. 

```{r echo=FALSE}
myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality.group)
colScale <- scale_colour_manual(name = "quality.group",values = myColors)

ggplot(aes(x = alcohol, y = volatile.acidity),
       data = wine) +
       geom_point(aes(color = quality.group)) +
       colScale
```

When using the grouped quality ratings variable, a trend is easier to see.
Higher rated wines dispersed across all abv levels, tend to lower volatile
acidity. Average tend to have lower alcohol levels and have more variance
across volatile acidity. I am surprised that I am not seeing low quality
wines grouped together. 

```{r echo=FALSE}
myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality)
colScale <- scale_colour_manual(name = "quality",values = myColors)

ggplot(aes(x = alcohol, y = citric.acid),
       data = wine) +
       geom_point(aes(color = quality)) +
       colScale
```

This chart with citric.acid on the y scale shows a lot of dispersion across
citric acid levels, yet again alcohol is grouped lower within a quality rating
of 5. By now we can start to see that most wines rated 5 in this data set,
regardless of other factors will mostly have alcohol levels inbetween 8-10 and
those rated 7 or 8 will mostly have abv of 10+. 


```{r echo=FALSE}
myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality)
colScale <- scale_colour_manual(name = "quality",values = myColors)

ggplot(aes(x = alcohol, y = residual.sugar),
       data = wine) +
       geom_point(aes(color = quality)) +
       colScale 
```


```{r echo=FALSE}
myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality)
colScale <- scale_colour_manual(name = "quality",values = myColors)

ggplot(aes(x = alcohol, y = residual.sugar),
       data = wine) +
       geom_point(aes(color = quality)) +
       colScale +
       scale_y_continuous(limits = c(1, 4))
```

I lost interest in residual sugar as a predictor early in my analysis, but I
wanted to see it next to alcohol, grouped by quality. I assume the wines with
higher residual sugar levels are desert wines or wines known for their sweeter
qualities. Zoomed in on where most of the data lies, there isn't much of a
trend to see other than reconfirming what has been seen with quality rating
trends in previous plots. 

```{r echo=FALSE}
myColors <- brewer.pal(6,"Set1")
names(myColors) <- levels(wine$quality)
colScale <- scale_colour_manual(name = "quality",values = myColors)

ggplot(aes(x = volatile.acidity, y = citric.acid),
       data = wine) +
       geom_point(aes(color = quality)) +
       colScale 
```

Intuitively, when we plot two measurements that deal with acidity, the 
closest thing to a linear relationship see among this run of plots emerges,
but with a lot of variance. There are groupings among quality ratings 5 and 7
that can be easily seen.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the
### investigation. Were there features that strengthened each other in terms
### of looking at your feature(s) of interest?

### &

### Were there any interesting or surprising interactions between features?

As alcohol seems to have the most significant impact on quality ratings, I
split it up into 3 groups: low, medium, and high alohol content so I could
see how these groups interacted with other variables of interest. 

Through these plots I noticed that of those wines in the group with high 
alcohol, none of these wines received a rating of 3. 

I was particularly interested in how volatile acidity would interact with these
groups as it seems to be a significant factor in wine quality ratings as well.
I noticed that the high alcohol group had lower acidity within all quality
ratings. Those wines that have high volatile acidity content also have many 
more quality ratings of 3. 

I also plotted citric acid among the 3 groups. This plot seems to show a
relationship with higher ratings and higher citric acid levels across all
groups. Again though, the high alcohol group was least affected by citric
acid levels vs quality ratings. Other than at quality rating 4, the high
alcohol group had less citric acid at every rating (5-8) vs the low and 
medium groups. 

Next I plotted suplphate levels by quality v quality rating by alcohol groups.
The relationship was very similar across all quality ratings other than 8. At
a rating of 8, it seems that wines with lower alcohol had higher levels of
suplphates. 

Throughout the scatterplots, the strongest trend visualized was that average 
rated wines tend to have lower alcohol and higher rated wines have higher
alcohol levels. 

### OPTIONAL: Did you create any models with your dataset?
### Discuss the strengths and limitations of your model.

Because the relationship between quality and alcohol are plotted within 6
quality ratings (3-8), I found it difficult to see if the relationship is
linear, although they don't look far from it.  

```{r echo=FALSE}
p1 <- ggplot(aes(x = quality, y = alcohol), data = wine) +
       geom_point(stat = 'summary', fun.y = median) 
       

p2 <- ggplot(aes(x = quality, y = citric.acid), data = wine) +
       geom_point(stat = 'summary', fun.y = median)

p3 <- ggplot(aes(x = quality, y = volatile.acidity), data = wine) +
       geom_point(stat = 'summary', fun.y = median)

p4 <- ggplot(aes(x = quality, y = sulphates), data = wine) +
       geom_point(stat = 'summary', fun.y = median)

grid.arrange(p1, p2, p3, p4, ncol=2)
```
 
In this group of charts, the points represent the median alcohol level among
each quality rating. There are relationships that seem to break out among
quality ratings, for example among quality ratings 6-8 within alcohol
measurements. Volatile acidity looks to have a linear relationship across
ratings and with it's Pearson correlation coefficient of -.39 might be a 
moderate predictor in a linear model. 

I attempted a simple model that test the relationship between quality and 4
other variables. Past alcohol and volatile acidity, the model doesn't appear
to predict much. With an r-squared value of .3, I do not see this model as 
dependably predictive.  

```{r echo=FALSE}
m1 <- lm(as.numeric(quality) ~ alcohol, data = wine)
m2 <- update(m1, ~ . + alcohol)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + sulphates)
mtable(m1, m2, m3, m4, m5)
```
------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x = quality.group, y = alcohol, fill = quality.group), 
       data = wine) +
  geom_boxplot() +
  ylab("Alcohol By Volume") +
  xlab("Quality Rating Group") +
  ggtitle("Distribution of ABV Within Quality Rating Groups") 

```
### Description One
This boxplot shows a clear difference between the three quality groups 
(low, average, high). Those wines have been rated a 7 or 8 in quality ratings
have a much higher mean of abv. There are some wines in the average group that
have high alcohol, shown as outliers on the boxplot, however this vizualization
shows that most wines that have a higher rating will have more alcohol than
those that are rated as low or average. 

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(x = as.factor(quality), y = volatile.acidity, fill = quality),
       data = wine)+
  geom_boxplot() +
  xlab("Quality Rating") +
  ylab("Volatile Acidity") +
  ggtitle("Distribution of Volatile Acidity by Quality Rating") 

```

### Description Two
This boxplot is a good respresentation of the negative correlation that
volatile acidity has with quality rating. It is clear that as lower rated 
wines have higher levels of volatile acidity, while higher rated wines have
lower volatile acidity. Those wines rated 7 or 8 have similar levels. It can
also be seen that among lower quality ratings, there more of a range among
volatile acidity measurements, but as quality ratings increase, the range
becomes more compact, with less outliers as well.  If you told me one wine
has a volatile acidity of .4 and the another .8, I would tell you I would 
like to drink the one with .4. 



### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(aes(x = quality, y = sulphates), 
       data = wine) +
       geom_point(aes(color = alcohol.group, shape = alcohol.group, size = 1), 
       stat = 'summary', fun.y = median) +
       xlab("Quality Rating") +
       ylab("Median Sulphate Levels") +
       ggtitle("Sulphates by Quality Rating within Alcohol Level Groups") +
       scale_size(guide = "none")
```

### Description Three
I found this plot interesting because it emerged that sulphate levels are quite
consistent across quality ratings among alcohol groups. No matter if a wine has
a low, medium, or high abv, among quality ratings of 4-7, wines follow a 
distinct pattern. Only at the rating of 8 is there divergence.  


# Reflection
This exercise gave me a better understanding of how trends and variables to
explore can emerge through visualization. I was able to use a correlation 
computation to find out which variables might be worth exploring, however 
it is the plots that help answer the subtleties of how one variable affects
another. I liked the idea of grouping properties of interest in this exercise
to help dig deeper into how the the property affected quality ratings. This 
project did reinforce the importance of running statistical test. I think 
visualizations of data can be exciting and lead you to believe that a 
particular variable is affecting another much more than it actually is. 
I honestly expected even my simple linear model to show higher r-squared values
than it did.  

If I were going to further explore this group I would want a lot more wines 
rated low (3 or lower) and high (8 and higher) so I could have more confidence
in my findings. 

```{r echo=FALSE}
table(wine$quality)
```

This data set only has 10 wines rated as a 3 and 18 rated as an 8. That makes
a good argument for splitting the quality ratings into 3 groups as I did to draw
conclusions, but one could be less confident about drawing conclusions from
observations within the wines rated as 3 or 8. Then again, using the groups
gives less of a chance to see unique trends that emerge among specific quality
ratings. 

